---
title: "rSeahorse-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rSeahorse-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rSeahorse)
```

## Description:
 
  - data: read from workbook "Rate Data"
  - event: read from workbook "Time Event"
  - data_min: transform the datetime to the minutes
  - data_mito: Calculation for different types of respiration
  - event_label: for A, B, C, D near the vertical line: annotation
  - well_label: for well name used in the quick viewing: geom_text
  
  - data_seahorse: 
      "data", "data_min", "event_label", "well_label",  "sheet_names"
  
  - well_exclude: list of outliers after quick viewing
  - bin_list: convert numerical minute column to categorical section column 
  
## Work flow

###  1. Import data
```{r}
file <-  system.file("extdata", "XFAssay.xls", package="rSeahorse" ) 
layout <- system.file("extdata", "layout.csv", package="rSeahorse" ) 

group_info <- rSeahorse::convert_layout_to_long(layout)

rsh <- import_seahorse(file = file, group_info = group_info)

lapply(rsh, head)


```

  
### 2. Quick View and exclude wells
```{r}

qview_seahorse(rsh$data_min, rsh$well_label)
wells_exclude <- c("6", "11", "15", "18", "20", "23")

```

### 3. Plot the main figure
```{r}

plot_seahorse(data_min = rsh$data_min,
              wells_exclude = wells_exclude,
              event_label = rsh$event_label,
              group_info = group_info,
              size_group = 6)

```
### 4. Calculate OCR subtype
```{r}
bin_list <- make_time_bin(rsh$event_label)


calculate_OCR_subtype(data_min = rsh$data_min,
                      bin_list = bin_list, 
                      return_type = "plot")

calculate_OCR_subtype(data_min = rsh$data_min,
                      bin_list = bin_list, 
                      return_type = "df")




```




```{r}




```

